---
title: "Project3_Background_Noise"
author: "Silvy van Kuijk & Tony Di Fiore"
date: "6/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Calculating Background Noise SPL levels for acoustic files

## Step 0: Load in Packages & Files

```{r}
library(tidyverse)
library(lubridate)
library(stringr)
library(tuneR)
library(seewave)
source('PAMGuide.R')

Data <- read_csv("https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/Project3_Final_Data_per_Recording.csv", col_names = TRUE)
```

## Step 1: Does the row in the .csv file contains a primate call with a start time of 1 minute or more after start time of recording?

```{r}
Data$StartTime <- str_split_fixed(Data$FileName, "_",3)[,3] #Pull the start time of the recording out of the file name.
Data$StartTime <- parse_date_time(Data$StartTime, "HMS")
Data$CallOnset <- parse_date_time(Data$CallOnset, "HMS")

Data$TimeDiff <- difftime(Data$CallOnset, Data$StartTime, units="secs")


```

## Step 2: If TimeDiff is >70 seconds, then find accompanying recording on external hard drive.
Files are stored here: D:/John_Blake_Data/Harpia 2013
Then, calculate broadband SPL in 60 seconds before onset of call (time in CallOnset column)

```{r}
Data$FileName <- paste0(Data$FileName, ".csv")

if (Data$TimeDiff > 70){
  readWave(filename, from = 1, to = 60, 
    units = c("seconds"), header = TRUE)
  
} 


```

## Step 3: Calculate background noise level (0-8 kHz and 0-2 kHz) and add it to the .csv file

```{r}

```

