---
title: "Project3_Data_Merging_and_Cleaning"
author: "Silvy van Kuijk & Tony Di Fiore"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Creating the complete Project 3 main data files needed for GLMM analysis.


```{r}
library(dplyr)
library(lubridate)
library(stringr)
library(tidyverse)
library(lunar)
```

#Step 1: Load in data files.

# 1.1 - Load in main data
```{r}
setwd('C:/Users/Silvy/Documents/R/Repos/Project3WeatherPatterns')
#Acoustic <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/Project3_Final_Data_per_Recording.csv"
Acoustic_Data <- read_csv("Project3_DataSheet_20210916.csv", col_names = TRUE)

Acoustic_Data$StartTime <- str_split_fixed(Acoustic_Data$FileName, "_",3)[,3]
Acoustic_Data$Date <- as.Date(mdy(Acoustic_Data$Date))


```

# 1.2 - Load in lunar phase data
```{r}
Lunar <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/Johnblake_dates_with_phase.csv" 
Lunar_Data <- read_csv("Johnblake_dates_with_phase_updated.csv", col_names = TRUE)
Lunar_Data$Date <- as.Date(mdy(Lunar_Data$Date))
Lunar_Data$LunarPhase <- lunar.phase(Lunar_Data$Date, name = TRUE)
```

# 1.3 - Load in temperature data - overnight averages
```{r}
Overnight <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/overnight_temperatures.csv"
Overnight_Temperatures <- read_csv("overnight_temperatures.csv", col_names = TRUE)
Overnight_Temperatures <- read_csv(Overnight, col_names = TRUE)
```

# 1.4 - Load in temperature data - hourly temperature and humidity
```{r}
Hourly <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/WeatherDataPerHour.csv"
Hourly_Weather_Data <- read_csv(Hourly, col_names = TRUE)
Hourly_Weather_Data <- read_csv("WeatherDataPerHour.csv", col_names = TRUE)
```

#Step 2: Join file 1.2 to file 1.1
```{r}
Acoustic_and_Lunar_Data <- left_join(Acoustic_Data, Lunar_Data, by = "Date")
```

# Step 3: Join file 1.3 to file 1.1
```{r}
colnames(Overnight_Temperatures)[colnames(Overnight_Temperatures) == 'today'] <- 'Date'
Overnight_Temperatures <- Overnight_Temperatures %>% 
  select(Date, today_average, lastnight_average, lastnight_min) 
Overnight_Temperatures$Date <- as.Date(mdy(Overnight_Temperatures$Date))

Acoustic_and_Lunar_and_Overnight <- left_join(Acoustic_and_Lunar_Data, Overnight_Temperatures, by = "Date")
```

# Step 4: Join file 1.4 to file 1.1 - Code Witten by Tony Di Fiore.
```{r}
Acoustic_and_Lunar_and_Overnight <- 
  Acoustic_and_Lunar_and_Overnight %>%
  mutate(Date2 = parse_date_time(Date, orders = c("Ymd")),
         hh = str_sub(`StartTime`,1,2),
         mm = str_sub(`StartTime`,3,4),
         ss = str_sub(`StartTime`,5,6),
         dt = make_datetime(year=year(Date2), month=month(Date2), day=day(Date2), hour = as.numeric(hh), min = as.numeric(mm), sec = as.numeric(ss))) %>%
  select(-c(Date2, hh, mm, ss)) %>%
  mutate(roundDateTime = round_date(dt, unit = "hour"))

Hourly_Weather_Data <- Hourly_Weather_Data %>%
  mutate(Date2 = parse_date_time(Date, orders = c("mdY")),
         Time2 = parse_date_time(Time, orders = c("HMS")),
         dt = make_datetime(year=year(Date2), month=month(Date2), day=day(Date2), hour = hour(Time2), min = minute(Time2), sec = second(Time2))) %>%
  select(-c(Date2, Time2)) %>%
  mutate(roundDateTime = round_date(dt, unit = "hour"))

Complete_Data <- left_join(Acoustic_and_Lunar_and_Overnight, Hourly_Weather_Data[ , c("roundDateTime", "Celsius", "Humidity")], by = "roundDateTime")
```

The first data set is now complete. This shows us all data available per recording. Next, we need a data file that summarized the data per day.

# Step 5: Splitting 'PrimateSpecies' column into a 'Titi' and 'Howler' column
```{r}
Complete_Data <- mutate(Complete_Data, Titis = ifelse(PrimateSpecies == "Titi", TRUE, FALSE))
Complete_Data <- mutate(Complete_Data, Howlers = ifelse(PrimateSpecies == "Howler", TRUE, FALSE))
```

# Step 6: Correct duration data in 'CallDuration' column from minutes and seconds since midnight to seconds.
```{r}
Complete_Data$CallDurationInSec <- minute(Complete_Data$CallDuration) * 60 + second(Complete_Data$CallDuration)
```

#Step 7: Create a Year variable
```{r}
Complete_Data$Year <- format(as.Date(Complete_Data$Date, format="%m/%d/%Y"),"%Y")

```

# Step 8: Let's create 2 separate datasets that summarize data per day for each of the species.
```{r}
HowlersInDay <- Complete_Data %>% filter(Howlers == TRUE) %>% group_by(Date) %>% dplyr::summarize(HowlerCount = sum(Howlers), TotalCallTime = sum(CallDurationInSec, na.rm = TRUE))
HowlersInDay <- left_join(Lunar_Data, HowlersInDay, by = "Date")
HowlersInDay$HowlerCount[is.na(HowlersInDay$HowlerCount)] <- 0
HowlersInDay$TotalCallTime[is.na(HowlersInDay$TotalCallTime)] <- 0
HowlersInDay <- left_join(HowlersInDay, Overnight_Temperatures, by = "Date")
HowlersInDay$Year <- format(as.Date(HowlersInDay$Date, format="%m/%d/%Y"),"%Y")

TitisInDay <- Complete_Data %>% filter(Titis == TRUE) %>% group_by(Date) %>% summarise(TitiCount = sum(Titis), TotalCallTime = sum(CallDurationInSec, na.rm = TRUE))
TitisInDay <- left_join(Lunar_Data, TitisInDay, by = "Date")

TitisInDay$TitiCount[is.na(TitisInDay$TitiCount)] <- 0
TitisInDay$TotalCallTime[is.na(TitisInDay$TotalCallTime)] <- 0

TitisInDay <- left_join(TitisInDay, Overnight_Temperatures, by = "Date")
TitisInDay$Year <- format(as.Date(TitisInDay$Date, format="%m/%d/%Y"),"%Y")
```


Write datafiles
```{r}
write.csv(Complete_Data,"Complete_Data_Per_File_20210917.csv",na="NA",row.names=TRUE)

write.csv(HowlersInDay,"Howler_Data_Per_Day_20210917.csv",na="NA",row.names=TRUE)

write.csv(TitisInDay,"Titi_Data_Per_Day_20210917.csv",na="NA",row.names=TRUE)

```


