---
title: "Project3_Data_Merging_and_Cleaning"
author: "Silvy van Kuijk & Tony Di Fiore"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Creating the complete Project 3 main data files needed for GLMM analysis.


```{r}
#setwd('C:/Users/Silvy/Documents/R/Repos/Project3WeatherPatterns')
library(dplyr)
library(lubridate)
library(stringr)
library(tidyverse)
```

#Step 1: Load in data files.

# 1.1 - Load in main data
```{r}
Acoustic <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/Project3-Main_Data_Sheet.csv"
Acoustic_Data <- read_csv(Acoustic, col_names = TRUE)
```

# 1.2 - Load in lunar phase data
```{r}
Lunar <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/Johnblake_dates_with_phase.csv" 
Lunar_Data <- read_csv(Lunar, col_names = TRUE)
```

# 1.3 - Load in temperature data - overnight averages
```{r}
Overnight <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/overnight_temperatures.csv"
Overnight_Temperatures <- read_csv(Overnight, col_names = TRUE)
```

# 1.4 - Load in temperature data - hourly temperature and humidity
```{r}
Hourly <- "https://raw.githubusercontent.com/Callicebus/Project3WeatherPatterns/main/WeatherDataPerHour.csv"
Hourly_Weather_Data <- read_csv(Hourly, col_names = TRUE)
```

#Step 2: Join file 1.2 to file 1.1
```{r}
Acoustic_and_Lunar_Data <- left_join(Acoustic_Data, Lunar_Data, by = "Date")
```

# Step 3: Join file 1.3 to file 1.1
```{r}
colnames(Overnight_Temperatures)[colnames(Overnight_Temperatures) == 'today'] <- 'Date'
Overnight_Temperatures <- Overnight_Temperatures %>% 
  select(Date, today_average, lastnight_average, lastnight_min) 
Acoustic_and_Lunar_and_Overnight <- left_join(Acoustic_and_Lunar_Data, Overnight_Temperatures, by = "Date")
```

# Step 4: Join file 1.4 to file 1.1 - Code Witten by Tony Di Fiore.
```{r}
#Using read_csv rather than read.csv, I now get incorrect date/times for roundDateTime with this code below.
Acoustic_and_Lunar_and_Overnight <- 
  Acoustic_and_Lunar_and_Overnight %>%
  mutate(Date2 = parse_date_time(Date, orders = c("mdy")),
         hh = str_sub(`StartTime`,1,2),
         mm = str_sub(`StartTime`,3,4),
         ss = str_sub(`StartTime`,5,6),
         dt = make_datetime(year=year(Date2), month=month(Date2), day=day(Date2), hour = as.numeric(hh), min = as.numeric(mm), sec = as.numeric(ss))) %>%
  select(-c(Date2, hh, mm, ss)) %>%
  mutate(roundDateTime = round_date(dt, unit = "hour"))

Hourly_Weather_Data <- Hourly_Weather_Data %>%
  mutate(Date2 = parse_date_time(Date, orders = c("mdy")),
         Time2 = parse_date_time(Time, orders = c("HMS")),
         dt = make_datetime(year=year(Date2), month=month(Date2), day=day(Date2), hour = hour(Time2), min = minute(Time2), sec = second(Time2))) %>%
  select(-c(Date2, Time2)) %>%
  mutate(roundDateTime = round_date(dt, unit = "hour"))

#Something goes wrong here somewhere along the line. Not all Temp and Humidity data gets joined correctly. Not sure why.
Complete_Data <- left_join(Acoustic_and_Lunar_and_Overnight, Hourly_Weather_Data[ , c("roundDateTime", "Celsius", "Humidity")], by = "roundDateTime")
```

# Step 5: Splitting 'PrimateSpecies' column into a 'Titi' and 'Howler' column
```{r}
Complete_Data <- mutate(Complete_Data, Titis = ifelse(PrimateSpecies == "Titi", TRUE, FALSE))
Complete_Data <- mutate(Complete_Data, Howlers = ifelse(PrimateSpecies == "Howler", TRUE, FALSE))
```

# Step 6: Correct duration data in 'CallDuration' column from minutes and seconds since midnight to seconds.
```{r}
Complete_Data$CallDurationInSec <- minute(Complete_Data$CallDuration) * 60 + second(Complete_Data$CallDuration)
```

# Step 7: Let's create 2 separate datasets that summarize data per day for each of the species.
```{r}
HowlersInDay <- Complete_Data %>% filter(Howlers == TRUE) %>% group_by(Date) %>% dplyr::summarize(HowlerCount = sum(Howlers), TotalCallTime = sum(CallDurationInSec, na.rm = TRUE))

HowlersInDay <- left_join(Lunar_Data, HowlersInDay, by = "Date")

HowlersInDay$HowlerCount[is.na(HowlersInDay$HowlerCount)] <- 0

HowlersInDay <- left_join(HowlersInDay, Overnight_Temperatures, by = "Date")
HowlersInDay$Year <- format(as.Date(HowlersInDay$Date, format="%m/%d/%Y"),"%Y")

TitisInDay <- Complete_Data %>% filter(Titis == TRUE) %>% group_by(Date) %>% summarise(titiCount = sum(Titis), totalCallTime = sum(CallDurationInSec, na.rm = TRUE))



#Dates_Lunar_Overnight <- semi_join(Dates_Lunar_Overnight, data[ , c("Titis")], by = "Date", copy = TRUE)
#write.csv(Dates_Lunar_Overnight, "ForMayFourteen.csv",na="NA",row.names=TRUE)


#write.csv(Complete_Date, "Project3_Final_Data_per_Recording.csv",na="NA",row.names=TRUE)

